---
title: Rat Sighting Data 
output: 
  html_document:
    toc: true
    toc_float: true
    favicon: "favicon.ico"
---

Below are the descriptive statistics for our datasets. We decided against running regression models for a few significant reasons: the outcome of rat sightings is not continuous and is technically classified as counts, which would prompt us to run a Poisson regression instead of a linear regression. We could not, however, reformat our rat sighting observations into counts for a Poisson regression model, so we agreed that descriptive statistics regarding rat sightings, median income, and population would be the most appropriate way for us to work with the data.

```{r, message=FALSE, echo=FALSE}
library(tidyverse)

rat_sightings_data <- read.csv("./datasets_clean/rat_sightings_data")
population_individual_data <- read.csv("./datasets_clean/population_individual_data")
population_totals_data <- read.csv("./datasets_clean/population_totals_data")
income_data_2020 <- read.csv("./datasets_clean/income_data_2020")
```


### Frequency Table

```{r, message=FALSE, echo=FALSE}
avg_rat_sightings_freq = 
  rat_sightings_data %>% 
  mutate(unique_key = as_factor(unique_key)) %>% 
  group_by(year_created, borough) %>% 
  summarize(count = n_distinct(unique_key)) %>% 
  rename(Year = year_created) %>% 
  pivot_wider(
    names_from = borough,
    values_from = count
  )

knitr::kable(avg_rat_sightings_freq)
```

The table shows the frequency of rat sightings in the five NYC boroughs by year, from 2015 to 2020. Rat sighting frequency generally is highest in Brooklyn and lowest in Staten Island. Rat sightings in all boroughs aside from Manhattan peak in 2017.

### Measures of Central Tendency

**Measures of Central Tendency for Rat Sightings by Year**

```{r, message=FALSE, echo=FALSE}
avg_rat_sightings_box = 
  rat_sightings_data %>% 
  mutate(unique_key = as_factor(unique_key)) %>% 
  group_by(year_created, borough) %>% 
  summarize(count = n_distinct(unique_key)) %>% 
  rename(year = year_created)

summary_stats_numeric = 
  avg_rat_sightings_box%>%
  summarise(
    mean = mean(count),
    median = median(count),
    sd = sd(count),
    min = min(count),
    max = max(count)
  )

knitr::kable(summary_stats_numeric)
```

This table depicts the measures of central tendency for rat sightings by year: the mean, median, standard deviation, minimum, and maximum for 2015 to 2020. All measures peak in the year 2017, and 2020 presents the lowest rat sighting values yet.

**Measures of Central Tendency for Rat Sightings by Borough (2015-2020)**

```{r, message=FALSE, echo=FALSE}
avg_rat_sightings_box = 
  rat_sightings_data %>% 
  mutate(unique_key = as_factor(unique_key)) %>% 
  group_by(year_created, borough) %>% 
  summarize(count = n_distinct(unique_key)) %>% 
  rename(year = year_created)

summary_stats_borough = 
  avg_rat_sightings_box%>%
  group_by(borough) %>% 
  summarise(
    mean = mean(count),
    median = median(count),
    sd = sd(count),
    min = min(count),
    max = max(count)
  )

knitr::kable(summary_stats_borough)
```

This table depicts the measures of central tendency for rat sightings by borough: the mean, median, standard deviation, minimum, and maximum for 2015 to 2020. Brooklyn presents the highest values for all measures, with Manhattan having the second-highest values. Staten Island has the lowest central tendency measures.

### Measures of Dispersion 

**Measures of Dispersion for Rat Sightings by Year**

```{r, message=FALSE, echo=FALSE}
avg_rat_sightings_box = 
  rat_sightings_data %>% 
  mutate(unique_key = as_factor(unique_key)) %>% 
  group_by(year_created, borough) %>% 
  summarize(count = n_distinct(unique_key)) %>% 
  rename(year = year_created)

dispersion_stats = 
  avg_rat_sightings_box %>%
  summarise(
    range = max(count) - min(count),
    iqr = IQR(count),
    sd = sd(count),
    var = var(count)
  )

knitr::kable(dispersion_stats)
```

This table shows the measures of dispersion for rat sightings by year, from 2015 to 2020. The range, standard deviation, and variance are greatest in 2017. The interquartile range is greatest in 2015 and lowest in 2017.

**Measures of Dispersion for Rat Sightings by Borough (2015-2020)**

```{r, message=FALSE, echo=FALSE}
avg_rat_sightings_box = 
  rat_sightings_data %>% 
  mutate(unique_key = as_factor(unique_key)) %>% 
  group_by(year_created, borough) %>% 
  summarize(count = n_distinct(unique_key)) %>% 
  rename(year = year_created)

summary_stats_borough = 
  avg_rat_sightings_box%>%
  group_by(borough) %>% 
  summarise(
    range = max(count) - min(count),
    iqr = IQR(count),
    sd = sd(count),
    var = var(count)
  )

knitr::kable(summary_stats_borough)
```

This table shows the measures of dispersion for rat sightings by borough, from 2015 to 2020. Brooklyn has the highest values for all measures (range, interquartile range, standard deviation, and variance). Staten Island has the lowest values for all measures. The Bronx has the second highest range.

### Quartiles

**Quartiles for Rat Sightings by Year**

```{r, message=FALSE, echo=FALSE}
avg_rat_sightings_box = 
  rat_sightings_data %>% 
  mutate(unique_key = as_factor(unique_key)) %>% 
  group_by(year_created, borough) %>% 
  summarize(count = n_distinct(unique_key)) %>% 
  rename(year = year_created)

quartiles_quantile =
avg_rat_sightings_box %>%
  summarise(
    Q1 = quantile(count, 0.25),
    median = median(count),
    Q3 = quantile(count, 0.75)
  )

knitr::kable(quartiles_quantile)
```

This table depicts the quartiles for rat sightings from 2015 to 2020. The first quartile, median, and third quartile all increase until 2017. The first quartile and median values decrease after 2017, but the third quartile values increase from 2017 to 2019 until decreasing in 2020.

**Quartiles for Rat Sightings by Borough (2015-2020)**

```{r, message=FALSE, echo=FALSE}
avg_rat_sightings_box = 
  rat_sightings_data %>% 
  mutate(unique_key = as_factor(unique_key)) %>% 
  group_by(year_created, borough) %>% 
  summarize(count = n_distinct(unique_key)) %>% 
  rename(year = year_created)

quartiles_quantile_borough = 
  avg_rat_sightings_box %>%
  group_by(borough) %>% 
  summarise(
    Q1 = quantile(count, 0.25),
    median = median(count),
    Q3 = quantile(count, 0.75)
  )

knitr::kable(quartiles_quantile_borough)
```

This table depicts the quartiles for rat sightings by borough from 2015 to 2020. Brooklyn has the highest values for all quartiles, and Staten Island has the lowest.