---
title: Income Data
output: 
  html_document:
    toc: true
    toc_float: true
    favicon: "favicon.ico"
---

```{r, message=FALSE, echo=FALSE}
library(tidyverse)

rat_sightings_data <- read.csv("./datasets_clean/rat_sightings_data")
population_individual_data <- read.csv("./datasets_clean/population_individual_data")
population_totals_data <- read.csv("./datasets_clean/population_totals_data")
income_data_2020 <- read.csv("./datasets_clean/income_data_2020")
```

Below are the descriptive statistics for our datasets. We decided against running regression models for a few significant reasons: the outcome of rat sightings is not continuous and is technically classified as counts, which would prompt us to run a Poisson regression instead of a linear regression. We could not, however, reformat our rat sighting observations into counts for a Poisson regression model, so we agreed that descriptive statistics regarding rat sightings, median income, and population would be the most appropriate way for us to work with the data.

### Frequency Table

**Highest Frequency Median Income by Borough in 2020**

<details>
  <summary>Code</summary>
```{r, eval=FALSE}
med_income_freq = 
  income_data_2020 %>% 
  group_by(borough, median_income) |> 
  summarize(count = n()) |> 
  arrange(borough, desc(count)) |> 
  slice_head(n = 1) |> 
  select(-count) |> 
  pivot_wider(
    names_from = borough,
    values_from = median_income
  )

knitr::kable(med_income_freq, align = "l")
```
</details>

```{r, message=FALSE, echo=FALSE}
med_income_freq = 
  income_data_2020 %>% 
  group_by(borough, median_income) |> 
  summarize(count = n()) |> 
  arrange(borough, desc(count)) |> 
  slice_head(n = 1) |> 
  select(-count) |> 
  pivot_wider(
    names_from = borough,
    values_from = median_income
  )

knitr::kable(med_income_freq, align = "l")
```

This table shows the median income that was reported most frequently by borough in the year of 2020. The highest median income value reported most frequently is 31,347, from Staten Island, and the lowest is 14,213, from the Bronx.


### Measures of Central Tendency

**Measures of Central Tendency for Median Income by Borough in 2020**

<details>
  <summary>Code</summary>
```{r, eval=FALSE}
central_tend_med_income = 
  income_data_2020 %>% 
  group_by(borough) |> 
  summarise(
    mean = mean(median_income),
    median = median(median_income),
    sd = sd(median_income),
    min = min(median_income),
    max = max(median_income)
  )

knitr::kable(central_tend_med_income, align = "l")
```
</details>

```{r, message=FALSE, echo=FALSE}
central_tend_med_income = 
  income_data_2020 %>% 
  group_by(borough) |> 
  summarise(
    mean = mean(median_income),
    median = median(median_income),
    sd = sd(median_income),
    min = min(median_income),
    max = max(median_income)
  )

knitr::kable(central_tend_med_income, align = "l")
```

This table depicts the measures of central tendency for median income by borough in 2020. Manhattan has the highest mean, median, standard deviation, and maximum. Staten Island, however, has the highest minimum reported median income.

### Measures of Dispersion

**Measures of Dispersion for Median Income by Borough in 2020**

<details>
  <summary>Code</summary>
```{r, eval=FALSE}
disp_med_income = 
  income_data_2020 %>%
  group_by(borough) %>% 
  summarise(
    range = max(median_income) - min(median_income),
    iqr = IQR(median_income),
    sd = sd(median_income),
    var = var(median_income)
  )

knitr::kable(disp_med_income, align = "l")
```
</details>

```{r, message=FALSE, echo=FALSE}
disp_med_income = 
  income_data_2020 %>%
  group_by(borough) %>% 
  summarise(
    range = max(median_income) - min(median_income),
    iqr = IQR(median_income),
    sd = sd(median_income),
    var = var(median_income)
  )

knitr::kable(disp_med_income, align = "l")
```

This table depicts the measures of dispersion for median income by borough in 2020. Manhattan has the largest range, interquartile range, standard deviation, and variance. Staten Island has the lowest values for all measures except for the interquartile range.


### Quartiles

**Quartiles for Median Income by Borough in 2020**

<details>
  <summary>Code</summary>
```{r, eval=FALSE}
quartiles_med_income = 
  income_data_2020 %>%
  group_by(borough) %>% 
  summarise(
    Q1 = quantile(median_income, 0.25),
    median = median(median_income),
    Q3 = quantile(median_income, 0.75)
  )

knitr::kable(quartiles_med_income, align = "l")
```
</details>

```{r, message=FALSE, echo=FALSE}
quartiles_med_income = 
  income_data_2020 %>%
  group_by(borough) %>% 
  summarise(
    Q1 = quantile(median_income, 0.25),
    median = median(median_income),
    Q3 = quantile(median_income, 0.75)
  )

knitr::kable(quartiles_med_income, align = "l")
```

This table shows the quartiles for median income by borough in 2020. The largest first quartile value is from Staten Island, but Manhattan has the largest median and third quartile values. The Bronx has the lowest quartile values.
