---
title: "Data Cleaning/Tidying"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Loading Necessary Packages 
```{r}
library(tidyverse)
```

# Income Data Cleaning

```{r message=FALSE}
income_data_2020 = 
  read_csv("./datasets/Neighborhood_Financial_Health_Digital_Mapping_and_Data_Tool_20231125.csv") %>% 
  janitor::clean_names() %>% 
  select(puma, borough, neighborhoods, nyc_poverty_rate, median_income, perc_white, perc_black, perc_asian, perc_other, perc_hispanic)
```

## Check for missing values and Typos in neighborhood names

```{r}
missing_val_check = 
  income_data_2020 %>% 
  summarise_all(~sum(is.na(.)))

income_data_2020 %>% 
  group_by(neighborhoods)%>% 
  summarise(count= n())
```

There are no missing values in the dataset. There are 7 of each neghborhood names indicating that there are no typos in the dataset for neighborhood name. 

## Exporting clean dataset

```{r}
write_csv(income_data_2020, "datasets_clean/income_data_2020")
```

# Rat Sighting Data Cleaning 

```{r message=FALSE, warning=FALSE}
rat_sightings_data = 
  read_csv("./datasets/Rat_Sightings_20231125.csv", skip=1) %>% 
  janitor::clean_names() %>% 
  select(unique_key, created_date, closed_date, location_type, incident_zip, borough, latitude, longitude) %>% 
   mutate(
    borough = ifelse(borough == "MANHATTAN", "Manhattan", borough),
    borough = ifelse(borough == "BRONX", "Bronx", borough),
    borough = ifelse(borough == "QUEENS", "Queens", borough),
    borough = ifelse(borough == "BROOKLYN", "Brooklyn", borough),
    borough = ifelse(borough == "STATEN ISLAND", "Staten Island", borough)) %>% 
  filter(borough != "Unspecified") %>% 
  separate(col = created_date, into = c("month_created", "day_created", "year_created", "time_created")) %>% 
  separate(col = closed_date, into = c("month_closed", "day_closed", "year_closed", "time_closed")) %>% 
  select(-time_created, -time_closed) %>% 
  filter(
    year_created %in% 2015:2020,
    year_closed %in% 2015:2020
  ) %>% 
  filter(incident_zip != 83)
```

## Check for missing values
```{r}
missing_val_check_rat = 
  rat_sightings_data %>% 
  summarise_all(~sum(is.na(.)))
```

There are 662 missing values for both longitude and latitude but we will keep these in our tidied data.


## Exporting clean dataset

```{r}
write_csv(rat_sightings_data, "datasets_clean/rat_sightings_data")
```

# Population Data Cleaning